---
title: "Aging Biomarkers"
author: "Petr Simecek"
output: html_document
---

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

```{r knitr-opts-chunk, include=FALSE}
```

**Last updated:** `r Sys.Date()`

**Code version:** `r workflowr::extract_commit(".", 1)$sha1`

If we take aging biomarkers (genes or proteins with expression associated with age), is the aging signature shared between Heart and Kidney datasets? 

In another words, do samples that look older than actual age look older in both tissues?

## C130026I21Rik

`C130026I21Rik` is the most Age-associated gene in kidney (and the 5th in heart). The aging signature seems to be shared between tissues (p<0.001). 

```{r, echo=FALSE, fig.height=6, fig.width=6, message=FALSE, warning=FALSE}
kidney = new.env()
heart = new.env()
load("../../kidney2/R/DO188b_kidney_noprobs.RData", env=kidney)
load("../../heart2/DO189_heart_v2_noprobs.RData", env=heart)

suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(ggsci))
theme_set(theme_bw())

pval.output <- function(p) {
  if (p<0.001) 
    return("p < 0.001") 
  else 
    paste("p =",round(p,3))
}

# for some reasons, heart SN are characters
heart$annot.samples$Sample.Number <- as.numeric(heart$annot.samples$Sample.Number)

sn <- intersect(kidney$annot.samples$Sample.Number, heart$annot.samples$Sample.Number)
kidney$idx <- match(sn, kidney$annot.samples$Sample.Number)
heart$idx <- match(sn, heart$annot.samples$Sample.Number)

sel.gene <- "C130026I21Rik"
sel.kidney <- which(kidney$annot.mrna$symbol == sel.gene)
sel.heart <- which(heart$annot.mrna$symbol == sel.gene)

sel.mrna <- cbind(kidney$annot.samples[kidney$idx, ], 
              Kidney = kidney$expr.mrna[kidney$idx, sel.kidney],
              Heart = heart$expr.mrna[heart$idx, sel.heart])
sel.mrna$Age <- factor(sel.mrna$Age, levels = c(6,12,18))

library(magrittr)
library(broom)
av <- lm(Kidney ~ Age + Heart, data=sel.mrna) %>%
  anova() %>%
  tidy() 
pval <- subset(av, term == "Heart")$p.value

ggplot(sel.mrna, aes(x = Kidney, y = Heart, color = Age, shape=Sex, group=Age)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE, size=0.2) +
  labs(title = "Protein Expression") +
  scale_color_uchicago() + 
  labs(title=sel.gene, subtitle = pval.output(pval))

```


## Cdkn2a

`Cdkn2a` is well-known aging signature. Unfortunatelly, it is not expressed in most samples and therefore it is hard to do between tissues comparison. 

```{r, echo=FALSE, fig.width=6, fig.height=6}
sel.gene <- "Cdkn2a"
sel.kidney <- which(kidney$annot.mrna$symbol == sel.gene)
sel.heart <- which(heart$annot.mrna$symbol == sel.gene)

sel.mrna <- cbind(kidney$annot.samples[kidney$idx, ], 
              Kidney = kidney$expr.mrna[kidney$idx, sel.kidney],
              Heart = heart$expr.mrna[heart$idx, sel.heart])
sel.mrna$Age <- factor(sel.mrna$Age, levels = c(6,12,18))

av <- lm(Kidney ~ Age + Heart, data=sel.mrna) %>%
  anova() %>%
  tidy() 
pval <- subset(av, term == "Heart")$p.value

ggplot(sel.mrna, aes(x = Kidney, y = Heart, color = Age, shape=Sex, group=Age)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE, size=0.2) +
  labs(title = "Protein Expression") +
  scale_color_uchicago() + 
  labs(title=sel.gene, subtitle =pval.output(pval))

```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```
