---
title: "Mediation examples"
author: "Petr Simecek"
output: html_document
---

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

```{r knitr-opts-chunk, include=FALSE}
```

**Last updated:** `r Sys.Date()`

**Code version:** `r workflowr::extract_commit(".", 1)$sha1`


A few selected patterns of Age/Sex effects on kidney/heart protein expression and its mediation by mRNA. The effect, that is mediated, is always visualised by a color, the other effect by a shape.

You can also use R/Shiny app: http://churchilldev.jax.org:49194/effectmed/ (JAX network only)

The report have been generated from R/Markdown file [examples.Rmd](https://github.com/simecek/TheAgingProteome/blob/master/analysis/examples.Rmd).

\ 

## Gsta1

glutathione S-transferase, alpha 1

```{r Gsta1, echo=FALSE, fig.width=10, fig.height=10}
# load data
kidney = new.env()
heart = new.env()
load("../../kidney2/R/DO188b_kidney_noprobs.RData", env=kidney)
load("../../heart2/DO189_heart_v2_noprobs.RData", env=heart)

suppressPackageStartupMessages(library(ggplot2))

gene.symbol <- "Gsta1"
tissue <- kidney

idx <- which(tissue$annot.protein$symbol == gene.symbol)
stopifnot(length(idx)==1)
dt <- cbind(tissue$annot.samples, 
            mrna = tissue$expr.mrna[,idx], 
            protein = tissue$expr.protein[,idx])
dt$Age <- factor(dt$Age)
ggplot(dt, aes(x=mrna, y=protein, color=Sex, shape=Age)) +
          geom_point(size=I(3)) +
          labs(title = gene.symbol, caption = "tissue: Kidney") +
          theme_light()
```

\ 

## Fmo5

flavin containing monooxygenase 5

```{r Fmo5, echo=FALSE, fig.width=10, fig.height=10}
gene.symbol <- "Fmo5"
tissue <- kidney

idx <- which(tissue$annot.protein$symbol == gene.symbol)
stopifnot(length(idx)==1)
dt <- cbind(tissue$annot.samples, 
            mrna = tissue$expr.mrna[,idx], 
            protein = tissue$expr.protein[,idx])
dt$Age <- factor(dt$Age)
ggplot(dt, aes(x=mrna, y=protein, color=Sex, shape=Age)) +
          geom_point(size=I(3)) +
          labs(title = gene.symbol, caption = "tissue: Kidney") +
          theme_light()
```

\ 

## Ace

angiotensin I converting enzyme (peptidyl-dipeptidase A) 1

```{r Ace, echo=FALSE, fig.width=10, fig.height=10}
gene.symbol <- "Ace"
tissue <- kidney

idx <- which(tissue$annot.protein$symbol == gene.symbol)
stopifnot(length(idx)==1)
dt <- cbind(tissue$annot.samples, 
            mrna = tissue$expr.mrna[,idx], 
            protein = tissue$expr.protein[,idx])
dt$Age <- factor(dt$Age)
ggplot(dt, aes(x=mrna, y=protein, color=Sex, shape=Age)) +
          geom_point(size=I(3)) +
          labs(title = gene.symbol, caption = "tissue: Kidney") +
          theme_light()
```

\ 

## Clca1

chloride channel accessory 1

```{r Clca1, echo=FALSE, fig.width=10, fig.height=10}
gene.symbol <- "Clca1"
tissue <- kidney

idx <- which(tissue$annot.protein$symbol == gene.symbol)
stopifnot(length(idx)==1)
dt <- cbind(tissue$annot.samples, 
            mrna = tissue$expr.mrna[,idx], 
            protein = tissue$expr.protein[,idx])
dt$Age <- factor(dt$Age)
ggplot(dt, aes(x=mrna, y=protein, color=Age, shape=Sex)) +
          geom_point(size=I(3)) +
          labs(title = gene.symbol, caption = "tissue: Kidney") +
          theme_light()
```

\ 

## Acta2

actin, alpha 2, smooth muscle, aorta

```{r Acta2, echo=FALSE, fig.width=10, fig.height=10}
gene.symbol <- "Acta2"
tissue <- kidney

idx <- which(tissue$annot.protein$symbol == gene.symbol)
stopifnot(length(idx)==1)
dt <- cbind(tissue$annot.samples, 
            mrna = tissue$expr.mrna[,idx], 
            protein = tissue$expr.protein[,idx])
dt$Age <- factor(dt$Age)
ggplot(dt, aes(x=mrna, y=protein, color=Age, shape=Sex)) +
          geom_point(size=I(3)) +
          labs(title = gene.symbol, caption = "tissue: Kidney") +
          theme_light()
```

\ 

## Slc1a4

solute carrier family 1 (glutamate/neutral amino acid transporter), member 4

```{r Slc1a4, echo=FALSE, fig.width=10, fig.height=10}
gene.symbol <- "Slc1a4"
tissue <- kidney

idx <- which(tissue$annot.protein$symbol == gene.symbol)
stopifnot(length(idx)==1)
dt <- cbind(tissue$annot.samples, 
            mrna = tissue$expr.mrna[,idx], 
            protein = tissue$expr.protein[,idx])
dt$Age <- factor(dt$Age)
ggplot(dt, aes(x=mrna, y=protein, color=Age, shape=Sex)) +
          geom_point(size=I(3)) +
          labs(title = gene.symbol, caption = "tissue: Kidney") +
          theme_light()
```

\ 

## Session Information

```{r session-info}
```
