---
title: "Examples Selected by Gary"
author: "Petr Simecek"
output: html_document
---

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

```{r knitr-opts-chunk, include=FALSE}
```

**Last updated:** `r Sys.Date()`

**Code version:** `r workflowr::extract_commit(".", 1)$sha1`

The report has been generated from R/Markdown file [examples-for-gary.Rmd](https://github.com/simecek/TheAgingProteome/blob/master/analysis/examples-for-gary.Rmd).

## Complex I Genes

### Age and Sex Effects 

In both tissues, Complex I protein expression decreases with Age and Sex differences disappear. For mRNA expression, Age effect is small or missing.

The list of Complex I genes was given me by Gary and can be found [here](https://github.com/simecek/TheAgingProteome/blob/master/data/MitoComplexIGenes.txt).

```{r, echo=FALSE, fig.height=4.5, fig.width=10, warning=FALSE}
# load data
kidney = new.env()
heart = new.env()
load("../../kidney2/R/DO188b_kidney_noprobs.RData", env=kidney)
load("../../heart2/DO189_heart_v2_noprobs.RData", env=heart)

suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(ggsci))
suppressPackageStartupMessages(library(gridExtra))
theme_set(theme_bw())

# read a list of Complex I genes
complex1 <- read.csv("../data/MitoComplexIGenes.txt", sep="\t", as.is=TRUE)

kidney$c1sel <- which(kidney$annot.mrna$symbol[1:kidney$N[["complete"]]] %in% complex1$Symbol)
kidney$pca.mrna <- prcomp(scale(as.matrix(kidney$expr.mrna)[,kidney$c1sel]))
kidney$pca.protein <- prcomp(scale(as.matrix(kidney$expr.protein)[,kidney$c1sel]))
kidney$pca1 <- cbind(kidney$annot.samples, 
                     mrna_pca1 = predict(kidney$pca.mrna)[,1],
                     prot_pca1 = predict(kidney$pca.protein)[,1])
kidney$pca1$Age <- factor(kidney$pca1$Age, levels = c(6,12,18))
p1 <- ggplot(kidney$pca1, aes(x = Age, y = prot_pca1, color = Sex, group=Sex)) +
  geom_jitter(width=0.1) +
  stat_summary(fun.y=mean, geom="line", size=1.2) +
  labs(title = "Protein Expression", subtitle = "Kidney") +
  scale_color_npg()
p2 <- ggplot(kidney$pca1, aes(x = Age, y = mrna_pca1, color = Sex, group=Sex)) +
  geom_jitter(width=0.1) +
  stat_summary(fun.y=mean, geom="line", size=1.2) +
  labs(title = "mRNA Expression", subtitle = "Kidney") +
  scale_color_npg()
grid.arrange(p1, p2, ncol = 2)
```

```{r, echo=FALSE, fig.width=10, fig.height=4.5}
heart$c1sel <- which(heart$annot.mrna$symbol[1:heart$N[["complete"]]] %in% complex1$Symbol)
heart$pca.mrna <- prcomp(scale(as.matrix(heart$expr.mrna)[,heart$c1sel]))
heart$pca.protein <- prcomp(scale(as.matrix(heart$expr.protein)[,heart$c1sel]))
heart$pca1 <- cbind(heart$annot.samples, 
                     mrna_pca1 = predict(heart$pca.mrna)[,1],
                     prot_pca1 = predict(heart$pca.protein)[,1])
heart$pca1$Age <- factor(heart$pca1$Age, levels = c(6,12,18))
p1 <- ggplot(heart$pca1, aes(x = Age, y = prot_pca1, color = Sex, group=Sex)) +
  geom_jitter(width=0.1) +
  stat_summary(fun.y=mean, geom="line", size=1.2) +
  labs(title = "Protein Expression", subtitle = "Heart") +
  scale_color_npg()
p2 <- ggplot(heart$pca1, aes(x = Age, y = mrna_pca1, color = Sex, group=Sex)) +
  geom_jitter(width=0.1) +
  stat_summary(fun.y=mean, geom="line", size=1.2) +
  labs(title = "mRNA Expression", subtitle = "Heart") +
  scale_color_npg()
grid.arrange(p1, p2, ncol = 2)
```

### Protein vs mRNA Expression
```{r, echo=FALSE, fig.width=10, fig.height=4.5}
p1 <- ggplot(kidney$pca1, aes(x = mrna_pca1, y = prot_pca1, color = Age, shape=Sex, group=Age)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE, size=0.2) +
  labs(title = "Kidney", subtitle = "mRNA vs. Protein") +
  scale_color_uchicago()
p2 <- ggplot(heart$pca1, aes(x = mrna_pca1, y = prot_pca1, color = Age, shape=Sex, group=Age)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE, size=0.2) +
  labs(title = "Heart", subtitle = "mRNA vs. Protein") +
  scale_color_uchicago()
grid.arrange(p1, p2, ncol = 2)
```

### Kidney vs Heart Expression
```{r, echo=FALSE, fig.width=10, fig.height=4.5}
# for some reasons, heart SN are characters
heart$annot.samples$Sample.Number <- as.numeric(heart$annot.samples$Sample.Number)

sn <- intersect(kidney$pca1$Sample.Number, heart$pca1$Sample.Number)
kidney$idx <- match(sn, kidney$pca1$Sample.Number)
heart$idx <- match(sn, heart$pca1$Sample.Number)

mrna <- cbind(kidney$annot.samples[kidney$idx, ], 
              Kidney = kidney$pca1$mrna_pca1[kidney$idx],
              Heart = heart$pca1$mrna_pca1[heart$idx])
mrna$Age <- factor(mrna$Age, levels = c(6,12,18))

prot <- cbind(kidney$annot.samples[kidney$idx, ], 
              Kidney = kidney$pca1$prot_pca1[kidney$idx],
              Heart = heart$pca1$prot_pca1[heart$idx])
prot$Age <- factor(prot$Age, levels = c(6,12,18))

p1 <- ggplot(prot, aes(x = Kidney, y = Heart, color = Age, shape=Sex, group=Age)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE, size=0.2) +
  labs(title = "Protein Expression", subtitle = "Kidney vs. Heart") +
  scale_color_uchicago()
p2 <- ggplot(mrna, aes(x = Kidney, y = Heart, color = Age, shape=Sex, group=Age)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE, size=0.2) +
  labs(title = "mRNA Expression", subtitle = "Kidney vs. Heart") +
  scale_color_uchicago()

grid.arrange(p1, p2, ncol = 2)
```

## Immune Infiltration

### Age and Sex Effects 

In both tissues, Ig protein expression increase with Age. For mRNA expression, kidney and heart differ.

I have used genes with "Ig" as the first two characters of MGI symbol as an approximation for the list of immune genes.

```{r, echo=FALSE, fig.height=4.5, fig.width=10}
# load data

# read genes starting with 
kidney$c1sel <- grep("^Ig", kidney$annot.mrna$symbol[1:kidney$N[["complete"]]])

kidney$pca.mrna <- prcomp(scale(as.matrix(kidney$expr.mrna)[,kidney$c1sel]))
kidney$pca.protein <- prcomp(scale(as.matrix(kidney$expr.protein)[,kidney$c1sel]))
kidney$pca1 <- cbind(kidney$annot.samples, 
                     mrna_pca1 = predict(kidney$pca.mrna)[,1],
                     prot_pca1 = predict(kidney$pca.protein)[,1])
kidney$pca1$Age <- factor(kidney$pca1$Age, levels = c(6,12,18))
p1 <- ggplot(kidney$pca1, aes(x = Age, y = prot_pca1, color = Sex, group=Sex)) +
  geom_jitter(width=0.1) +
  stat_summary(fun.y=mean, geom="line", size=1.2) +
  labs(title = "Protein Expression", subtitle = "Kidney") +
  scale_color_npg()
p2 <- ggplot(kidney$pca1, aes(x = Age, y = mrna_pca1, color = Sex, group=Sex)) +
  geom_jitter(width=0.1) +
  stat_summary(fun.y=mean, geom="line", size=1.2) +
  labs(title = "mRNA Expression", subtitle = "Kidney") +
  scale_color_npg()
grid.arrange(p1, p2, ncol = 2)
```

```{r, echo=FALSE, fig.width=10, fig.height=4.5}
heart$c1sel <- grep("^Ig", heart$annot.mrna$symbol[1:heart$N[["complete"]]])

heart$pca.mrna <- prcomp(scale(as.matrix(heart$expr.mrna)[,heart$c1sel]))
heart$pca.protein <- prcomp(scale(as.matrix(heart$expr.protein)[,heart$c1sel]))
heart$pca1 <- cbind(heart$annot.samples, 
                     mrna_pca1 = predict(heart$pca.mrna)[,1],
                     prot_pca1 = predict(heart$pca.protein)[,1])
heart$pca1$Age <- factor(heart$pca1$Age, levels = c(6,12,18))
p1 <- ggplot(heart$pca1, aes(x = Age, y = prot_pca1, color = Sex, group=Sex)) +
  geom_jitter(width=0.1) +
  stat_summary(fun.y=mean, geom="line", size=1.2) +
  labs(title = "Protein Expression", subtitle = "Heart") +
  scale_color_npg()
p2 <- ggplot(heart$pca1, aes(x = Age, y = mrna_pca1, color = Sex, group=Sex)) +
  geom_jitter(width=0.1) +
  stat_summary(fun.y=mean, geom="line", size=1.2) +
  labs(title = "mRNA Expression", subtitle = "Heart") +
  scale_color_npg()
grid.arrange(p1, p2, ncol = 2)
```

### Protein vs mRNA Expression
```{r, echo=FALSE, fig.width=10, fig.height=4.5}
p1 <- ggplot(kidney$pca1, aes(x = mrna_pca1, y = prot_pca1, color = Age, shape=Sex, group=Age)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE, size=0.2) +
  labs(title = "Kidney", subtitle = "mRNA vs. Protein") +
  scale_color_uchicago()
p2 <- ggplot(heart$pca1, aes(x = mrna_pca1, y = prot_pca1, color = Age, shape=Sex, group=Age)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE, size=0.2) +
  labs(title = "Heart", subtitle = "mRNA vs. Protein") +
  scale_color_uchicago()
grid.arrange(p1, p2, ncol = 2)
```

### Kidney vs Heart Expression
```{r, echo=FALSE, fig.width=10, fig.height=4.5}
# for some reasons, heart SN are characters
heart$annot.samples$Sample.Number <- as.numeric(heart$annot.samples$Sample.Number)

sn <- intersect(kidney$pca1$Sample.Number, heart$pca1$Sample.Number)
kidney$idx <- match(sn, kidney$pca1$Sample.Number)
heart$idx <- match(sn, heart$pca1$Sample.Number)

mrna <- cbind(kidney$annot.samples[kidney$idx, ], 
              Kidney = kidney$pca1$mrna_pca1[kidney$idx],
              Heart = heart$pca1$mrna_pca1[heart$idx])
mrna$Age <- factor(mrna$Age, levels = c(6,12,18))

prot <- cbind(kidney$annot.samples[kidney$idx, ], 
              Kidney = kidney$pca1$prot_pca1[kidney$idx],
              Heart = heart$pca1$prot_pca1[heart$idx])
prot$Age <- factor(prot$Age, levels = c(6,12,18))

p1 <- ggplot(prot, aes(x = Kidney, y = Heart, color = Age, shape=Sex, group=Age)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE, size=0.2) +
  labs(title = "Protein Expression", subtitle = "Kidney vs. Heart") +
  scale_color_uchicago()
p2 <- ggplot(mrna, aes(x = Kidney, y = Heart, color = Age, shape=Sex, group=Age)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE, size=0.2) +
  labs(title = "mRNA Expression", subtitle = "Kidney vs. Heart") +
  scale_color_uchicago()

grid.arrange(p1, p2, ncol = 2)
```

## Myosins

### Age and Sex Effects 

For kidney protein expression, Myosins seem to be clearly associated to Age. This pattern does not to be seens in other situations.

I have used genes with "Myo" as the first three characters of MGI symbol as an approximation for the list of myosins.

```{r, echo=FALSE, fig.height=4.5, fig.width=10}
# load data

# read genes starting with 
kidney$c1sel <- grep("^Myo", kidney$annot.mrna$symbol[1:kidney$N[["complete"]]])

kidney$pca.mrna <- prcomp(scale(as.matrix(kidney$expr.mrna)[,kidney$c1sel]))
kidney$pca.protein <- prcomp(scale(as.matrix(kidney$expr.protein)[,kidney$c1sel]))
kidney$pca1 <- cbind(kidney$annot.samples, 
                     mrna_pca1 = predict(kidney$pca.mrna)[,1],
                     prot_pca1 = predict(kidney$pca.protein)[,1])
kidney$pca1$Age <- factor(kidney$pca1$Age, levels = c(6,12,18))
p1 <- ggplot(kidney$pca1, aes(x = Age, y = prot_pca1, color = Sex, group=Sex)) +
  geom_jitter(width=0.1) +
  stat_summary(fun.y=mean, geom="line", size=1.2) +
  labs(title = "Protein Expression", subtitle = "Kidney") +
  scale_color_npg()
p2 <- ggplot(kidney$pca1, aes(x = Age, y = mrna_pca1, color = Sex, group=Sex)) +
  geom_jitter(width=0.1) +
  stat_summary(fun.y=mean, geom="line", size=1.2) +
  labs(title = "mRNA Expression", subtitle = "Kidney") +
  scale_color_npg()
grid.arrange(p1, p2, ncol = 2)
```

```{r, echo=FALSE, fig.width=10, fig.height=4.5}
heart$c1sel <- grep("^Myo", heart$annot.mrna$symbol[1:heart$N[["complete"]]])

heart$pca.mrna <- prcomp(scale(as.matrix(heart$expr.mrna)[,heart$c1sel]))
heart$pca.protein <- prcomp(scale(as.matrix(heart$expr.protein)[,heart$c1sel]))
heart$pca1 <- cbind(heart$annot.samples, 
                     mrna_pca1 = predict(heart$pca.mrna)[,1],
                     prot_pca1 = predict(heart$pca.protein)[,1])
heart$pca1$Age <- factor(heart$pca1$Age, levels = c(6,12,18))
p1 <- ggplot(heart$pca1, aes(x = Age, y = prot_pca1, color = Sex, group=Sex)) +
  geom_jitter(width=0.1) +
  stat_summary(fun.y=mean, geom="line", size=1.2) +
  labs(title = "Protein Expression", subtitle = "Heart") +
  scale_color_npg()
p2 <- ggplot(heart$pca1, aes(x = Age, y = mrna_pca1, color = Sex, group=Sex)) +
  geom_jitter(width=0.1) +
  stat_summary(fun.y=mean, geom="line", size=1.2) +
  labs(title = "mRNA Expression", subtitle = "Heart") +
  scale_color_npg()
grid.arrange(p1, p2, ncol = 2)
```

### Protein vs mRNA Expression
```{r, echo=FALSE, fig.width=10, fig.height=4.5}
p1 <- ggplot(kidney$pca1, aes(x = mrna_pca1, y = prot_pca1, color = Age, shape=Sex, group=Age)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE, size=0.2) +
  labs(title = "Kidney", subtitle = "mRNA vs. Protein") +
  scale_color_uchicago()
p2 <- ggplot(heart$pca1, aes(x = mrna_pca1, y = prot_pca1, color = Age, shape=Sex, group=Age)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE, size=0.2) +
  labs(title = "Heart", subtitle = "mRNA vs. Protein") +
  scale_color_uchicago()
grid.arrange(p1, p2, ncol = 2)
```

### Kidney vs Heart Expression
```{r, echo=FALSE, fig.width=10, fig.height=4.5}
# for some reasons, heart SN are characters
heart$annot.samples$Sample.Number <- as.numeric(heart$annot.samples$Sample.Number)

sn <- intersect(kidney$pca1$Sample.Number, heart$pca1$Sample.Number)
kidney$idx <- match(sn, kidney$pca1$Sample.Number)
heart$idx <- match(sn, heart$pca1$Sample.Number)

mrna <- cbind(kidney$annot.samples[kidney$idx, ], 
              Kidney = kidney$pca1$mrna_pca1[kidney$idx],
              Heart = heart$pca1$mrna_pca1[heart$idx])
mrna$Age <- factor(mrna$Age, levels = c(6,12,18))

prot <- cbind(kidney$annot.samples[kidney$idx, ], 
              Kidney = kidney$pca1$prot_pca1[kidney$idx],
              Heart = heart$pca1$prot_pca1[heart$idx])
prot$Age <- factor(prot$Age, levels = c(6,12,18))

p1 <- ggplot(prot, aes(x = Kidney, y = Heart, color = Age, shape=Sex, group=Age)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE, size=0.2) +
  labs(title = "Protein Expression", subtitle = "Kidney vs. Heart") +
  scale_color_uchicago()
p2 <- ggplot(mrna, aes(x = Kidney, y = Heart, color = Age, shape=Sex, group=Age)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE, size=0.2) +
  labs(title = "mRNA Expression", subtitle = "Kidney vs. Heart") +
  scale_color_uchicago()

grid.arrange(p1, p2, ncol = 2)
```

## Actn4

Protein expression of *Actn4* increases with Age in both tissues. This effect cannot be seen on mRNA level. 

### Age and Sex Effects 


```{r, echo=FALSE, fig.height=4.5, fig.width=10}
# load data

# read genes starting with 
kidney$c1sel <- grep("Actn4", kidney$annot.mrna$symbol[1:kidney$N[["complete"]]])

kidney$pca.mrna <- prcomp(scale(as.matrix(kidney$expr.mrna)[,kidney$c1sel]))
kidney$pca.protein <- prcomp(scale(as.matrix(kidney$expr.protein)[,kidney$c1sel]))
kidney$pca1 <- cbind(kidney$annot.samples, 
                     mrna_pca1 = predict(kidney$pca.mrna)[,1],
                     prot_pca1 = predict(kidney$pca.protein)[,1])
kidney$pca1$Age <- factor(kidney$pca1$Age, levels = c(6,12,18))
p1 <- ggplot(kidney$pca1, aes(x = Age, y = prot_pca1, color = Sex, group=Sex)) +
  geom_jitter(width=0.1) +
  stat_summary(fun.y=mean, geom="line", size=1.2) +
  labs(title = "Protein Expression", subtitle = "Kidney") +
  scale_color_npg()
p2 <- ggplot(kidney$pca1, aes(x = Age, y = mrna_pca1, color = Sex, group=Sex)) +
  geom_jitter(width=0.1) +
  stat_summary(fun.y=mean, geom="line", size=1.2) +
  labs(title = "mRNA Expression", subtitle = "Kidney") +
  scale_color_npg()
grid.arrange(p1, p2, ncol = 2)
```

```{r, echo=FALSE, fig.width=10, fig.height=4.5}
heart$c1sel <- grep("Actn4", heart$annot.mrna$symbol[1:heart$N[["complete"]]])

heart$pca.mrna <- prcomp(scale(as.matrix(heart$expr.mrna)[,heart$c1sel]))
heart$pca.protein <- prcomp(scale(as.matrix(heart$expr.protein)[,heart$c1sel]))
heart$pca1 <- cbind(heart$annot.samples, 
                     mrna_pca1 = predict(heart$pca.mrna)[,1],
                     prot_pca1 = predict(heart$pca.protein)[,1])
heart$pca1$Age <- factor(heart$pca1$Age, levels = c(6,12,18))
p1 <- ggplot(heart$pca1, aes(x = Age, y = prot_pca1, color = Sex, group=Sex)) +
  geom_jitter(width=0.1) +
  stat_summary(fun.y=mean, geom="line", size=1.2) +
  labs(title = "Protein Expression", subtitle = "Heart") +
  scale_color_npg()
p2 <- ggplot(heart$pca1, aes(x = Age, y = mrna_pca1, color = Sex, group=Sex)) +
  geom_jitter(width=0.1) +
  stat_summary(fun.y=mean, geom="line", size=1.2) +
  labs(title = "mRNA Expression", subtitle = "Heart") +
  scale_color_npg()
grid.arrange(p1, p2, ncol = 2)
```

### Protein vs mRNA Expression
```{r, echo=FALSE, fig.width=10, fig.height=4.5}
p1 <- ggplot(kidney$pca1, aes(x = mrna_pca1, y = prot_pca1, color = Age, shape=Sex, group=Age)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE, size=0.2) +
  labs(title = "Kidney", subtitle = "mRNA vs. Protein") +
  scale_color_uchicago()
p2 <- ggplot(heart$pca1, aes(x = mrna_pca1, y = prot_pca1, color = Age, shape=Sex, group=Age)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE, size=0.2) +
  labs(title = "Heart", subtitle = "mRNA vs. Protein") +
  scale_color_uchicago()
grid.arrange(p1, p2, ncol = 2)
```

### Kidney vs Heart Expression
```{r, echo=FALSE, fig.width=10, fig.height=4.5}
# for some reasons, heart SN are characters
heart$annot.samples$Sample.Number <- as.numeric(heart$annot.samples$Sample.Number)

sn <- intersect(kidney$pca1$Sample.Number, heart$pca1$Sample.Number)
kidney$idx <- match(sn, kidney$pca1$Sample.Number)
heart$idx <- match(sn, heart$pca1$Sample.Number)

mrna <- cbind(kidney$annot.samples[kidney$idx, ], 
              Kidney = kidney$pca1$mrna_pca1[kidney$idx],
              Heart = heart$pca1$mrna_pca1[heart$idx])
mrna$Age <- factor(mrna$Age, levels = c(6,12,18))

prot <- cbind(kidney$annot.samples[kidney$idx, ], 
              Kidney = kidney$pca1$prot_pca1[kidney$idx],
              Heart = heart$pca1$prot_pca1[heart$idx])
prot$Age <- factor(prot$Age, levels = c(6,12,18))

p1 <- ggplot(prot, aes(x = Kidney, y = Heart, color = Age, shape=Sex, group=Age)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE, size=0.2) +
  labs(title = "Protein Expression", subtitle = "Kidney vs. Heart") +
  scale_color_uchicago()
p2 <- ggplot(mrna, aes(x = Kidney, y = Heart, color = Age, shape=Sex, group=Age)) +
  geom_point() +
  geom_smooth(method="lm", se=FALSE, size=0.2) +
  labs(title = "mRNA Expression", subtitle = "Kidney vs. Heart") +
  scale_color_uchicago()

grid.arrange(p1, p2, ncol = 2)
```

## Session information

<!-- Insert the session information into the document -->
```{r session-info}
```
