---
title: "Direction of Age Effect"
author: "Petr Simecek"
output: html_document
---

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

```{r knitr-opts-chunk, include=FALSE}
```

**Last updated:** `r Sys.Date()`

**Code version:** `r workflowr::extract_commit(".", 1)$sha1`

#### Gary's Question

*can you quickly assess how many of the age-associated changes in kidney&heart protein are increasing versus decreasing? The more i learn about mass-spec, the weirder it gets.  Depending on your reply - I will explain what I think might be going on.*

#### Short Answer

Strongly Age-associated proteins are more likely to be decreasing with Age than increasing with Age.

#### Long Answer

If percentage of down / up regulation is plotted against categorized p-value of dropping Age from the model, then for the most Age-associated proteins down regulation is clearly over 50 percent.

The report has been generated from R/Markdown file [increasing_vs_decreasing_with_age.Rmd](https://github.com/simecek/TheAgingProteome/blob/master/analysis/increasing_vs_decreasing_with_age.Rmd).

```{r cars, echo=FALSE}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))

kidney <- read.csv("../../kidney2/kidney_anova_table.csv", as.is=TRUE)
heart <- read.csv("../../heart2/heart_anova_table.csv", as.is=TRUE)

# transform summary statistics for kidney
kidney.summary <- kidney %>% 
  select(p.Prot_Age.Sex, average.Prot.Age.6, average.Prot.Age.18) %>%
  mutate(age_change = average.Prot.Age.18 - average.Prot.Age.6,
         age_direction = factor(ifelse(age_change<0,"down","up"), levels=c("up","down")),
         p.adj = p.adjust(p.Prot_Age.Sex, method="BH"),
         minus.log.p.adj = -log10(p.adj),
         minus.log.p.adj.cat = ifelse(round(minus.log.p.adj)<9,
                                round(minus.log.p.adj), "9+" )) 

# count ups and downs per -logP category 
tab.kidney <- kidney.summary %>%
  group_by(minus.log.p.adj.cat) %>%
  summarize(n=n())

# make a plot
ggplot(kidney.summary, aes(x=minus.log.p.adj.cat)) + 
  geom_bar(aes(fill=age_direction), position="fill") +
  geom_text(data = tab.kidney, aes(x=minus.log.p.adj.cat, label=n, y=1), vjust=-0.2) +
  geom_hline(yintercept=0.5, color="grey70") +
  theme_minimal() +
  labs(title = "Kidney",
       subtitle = "Number of proteins in each category is stated above the bars",
       y = "Up vs. Down",
       x = expression(-log[10](p-adjusted)),
       fill = "Direction of Age-Effect")

# transform summary statistics for heart
heart.summary <- heart %>% 
  select(p.Prot_Age.Sex, average.Prot.Age.6, average.Prot.Age.18) %>%
  mutate(age_change = average.Prot.Age.18 - average.Prot.Age.6,
         age_direction = factor(ifelse(age_change<0,"down","up"), levels=c("up","down")),
         p.adj = p.adjust(p.Prot_Age.Sex, method="BH"),
         minus.log.p.adj = -log10(p.adj),
         minus.log.p.adj.cat = ifelse(round(minus.log.p.adj)<9,
                                round(minus.log.p.adj), "9+" )) 

# count ups and downs per -logP category 
tab.heart <- heart.summary %>%
  group_by(minus.log.p.adj.cat) %>%
  summarize(n=n())

# make a plot
ggplot(heart.summary, aes(x=minus.log.p.adj.cat)) + 
  geom_bar(aes(fill=age_direction), position="fill") +
  geom_text(data = tab.heart, aes(x=minus.log.p.adj.cat, label=n, y=1), vjust=-0.2) +
  geom_hline(yintercept=0.5, color="grey70") +
  theme_minimal() +
  labs(title = "Heart",
       subtitle = "Number of proteins in each category is stated above the bars",
       y = "Up vs. Down",
       x = expression(-log[10](p-adjusted)),
       fill = "Direction of Age-Effect")
```


## Session Information

```{r session-info}
```
