---
title: "PCA plots"
author: "Petr Simecek"
output: html_document
---

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

```{r knitr-opts-chunk, include=FALSE}
```

**Last updated:** `r Sys.Date()`

**Code version:** `r workflowr::extract_commit(".", 1)$sha1`

\ 

PCA plots of gene/protein expression for kidney/heart tissue. We expect to see what are the main factors driving the expression and whether we have any outlying samples.

All plots below are interactive (tooltip = mouse id) and have been generated from R/Markdown file [pca.Rmd](https://github.com/simecek/TheAgingProteome/blob/master/analysis/pca.Rmd).

\ 

## Kidney mRNA-level PCA

```{r kidney_mrna, echo=FALSE}
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(plotly))

# load data - kidney
dt <- read.csv("../../kidney2/kidney_pca_predict.csv", as.is=TRUE)
dt$Age <- factor(dt$Age)
dt$text <- paste0(dt$Sex, dt$Sample.Number)

pl <- ggplot(dt, aes(x=PC1_mrna, y=PC2_mrna, shape=Sex, color=Age, text=text)) +   
  geom_point(size=I(3)) + 
  ggtitle("Kidney mRNA expression: PCA1 vs. PCA2") +
  theme_bw()
ggplotly(pl, tooltip="text")
```

\ 

## Kidney protein-level PCA

```{r kidney_protein, echo=FALSE}
pl<- ggplot(dt, aes(x=PC1_prot, y=PC2_prot, shape=Sex, color=Age, text=text)) +
  geom_point(size=I(3)) + 
  ggtitle("Kidney protein expression: PCA1 vs. PCA2") +
  theme_bw()
ggplotly(pl, tooltip="text")
```

\ 

## Heart mRNA-level PCA

```{r heart_mrna, echo=FALSE}
# load data - heart
dt <- read.csv("../../heart2/heart_pca_predict.csv", as.is=TRUE)
dt$Age <- factor(dt$Age)
dt$text <- paste0(dt$Sex, dt$Sample.Number)

pl <- ggplot(dt, aes(x=PC1_mrna, y=PC2_mrna, shape=Sex, color=Age, text=text)) +   
  geom_point(size=I(3)) + 
  ggtitle("Heart mRNA expression: PCA1 vs. PCA2") +
  theme_bw()
ggplotly(pl, tooltip="text")
```

\ 

## Heart protein-level PCA

```{r heart_protein, echo=FALSE}
pl <- ggplot(dt, aes(x=PC1_prot, y=PC2_prot, shape=Sex, color=Age, text=text)) +   
  geom_point(size=I(3)) + 
  ggtitle("Heart protein expression: PCA1 vs. PCA2") +
  theme_bw()
ggplotly(pl, tooltip="text")
```

\ 

## Session Information

```{r session-info}
```
