---
title: "PCA Contrast And Compare"
author: "Petr Simecek"
output: html_document
---

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

```{r knitr-opts-chunk, include=FALSE}
```

```{r data_load, echo=FALSE}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(corrplot))

# kidney data
pca.kidney <- read.csv("../../kidney2/kidney_pca_predict.csv", as.is=TRUE)
pca.kidney.mrna <- read.csv("../../kidney2/kidney_pca_predict_mrna.csv", as.is=TRUE)
pca.kidney.joined <- inner_join(pca.kidney,
                      pca.kidney.mrna, 
                  by = c("Mouse.ID","Sex","Age"),
                  suffix = c("_subset", "_all"))
# heart data
pca.heart <- read.csv("../../heart2/heart_pca_predict.csv", as.is=TRUE)
pca.heart.mrna <- read.csv("../../heart2/heart_pca_predict_mrna.csv", as.is=TRUE)
pca.heart.joined <- inner_join(pca.heart,
                      pca.heart.mrna, 
                  by = c("Mouse.ID","Sex","Age"),
                  suffix = c("_subset", "_all"))
# join of kidney & heart data into one table
pca.both <- inner_join(pca.kidney, pca.heart, 
                  by = "Mouse.ID",
                  suffix = c("_kidney", "_heart"))

# for ploting
pca.both$Age <- pca.both$Age_kidney
pca.both$Sex <- as.numeric(factor(pca.both$Sex_kidney))
```

**Last updated:** `r Sys.Date()`

**Code version:** `r workflowr::extract_commit(".", 1)$sha1`

Three PCA comparisons: tissue differences (kidney vs. heart), level differences (mRNA vs. protein) and subsetting differences (all vs. only genes with protein).

The report has been generated from R/Markdown file [pca_comparison.Rmd](https://github.com/simecek/TheAgingProteome/blob/master/analysis/pca_comparison.Rmd).

## Kidney vs. Heart

### mRNA-level

```{r kh_mrna, echo=FALSE}
sex.age <- grep("^(Sex|Age)$", names(pca.both))
set1 <- c(sex.age, grep("mrna_kidney$", names(pca.both)))
set2 <- c(sex.age, grep("mrna_heart$", names(pca.both)))

tmpcor <- cor(pca.both[,set1],pca.both[,set2])
corrplot(tmpcor)
```

### Protein-level

```{r kh_prot, echo=FALSE}
sex.age <- grep("^(Sex|Age)$", names(pca.both))
set1 <- c(sex.age, grep("prot_kidney$", names(pca.both)))
set2 <- c(sex.age, grep("prot_heart$", names(pca.both)))

tmpcor <- cor(pca.both[,set1],pca.both[,set2])
corrplot(tmpcor)
```

## mRNA vs Protein

### Kidney

```{r mp_kidney, echo=FALSE}
pca.kidney$Sex <- as.numeric(factor(pca.kidney$Sex))
sex.age <- grep("^(Sex|Age)$", names(pca.kidney))
set1 <- c(sex.age, grep("mrna$", names(pca.kidney)))
set2 <- c(sex.age, grep("prot$", names(pca.kidney)))

tmpcor <- cor(pca.kidney[,set1],pca.kidney[,set2])
corrplot(tmpcor)
```

### Heart

```{r mp_heart, echo=FALSE}
pca.heart$Sex <- as.numeric(factor(pca.heart$Sex))
sex.age <- grep("^(Sex|Age)$", names(pca.heart))
set1 <- c(sex.age, grep("mrna$", names(pca.heart)))
set2 <- c(sex.age, grep("prot$", names(pca.heart)))

tmpcor <- cor(pca.heart[,set1],pca.heart[,set2])
corrplot(tmpcor)
```

## All genes vs. genes with protein only

### Kidney

```{r full_kidney, echo=FALSE}
pca.kidney.joined$Sex <- as.numeric(factor(pca.kidney.joined$Sex))
sex.age <- grep("^(Sex|Age)$", names(pca.kidney.joined))
set1 <- c(sex.age, grep("mrna_all$", names(pca.kidney.joined)))
set2 <- c(sex.age, grep("mrna_subset$", names(pca.kidney.joined)))

tmpcor <- cor(pca.kidney.joined[,set1],pca.kidney.joined[,set2])
corrplot(tmpcor)
```

### Heart

```{r full_heart, echo=FALSE}
pca.heart.joined$Sex <- as.numeric(factor(pca.heart.joined$Sex))
sex.age <- grep("^(Sex|Age)$", names(pca.heart.joined))
set1 <- c(sex.age, grep("mrna_all$", names(pca.heart.joined)))
set2 <- c(sex.age, grep("mrna_subset$", names(pca.heart.joined)))

tmpcor <- cor(pca.heart.joined[,set1],pca.heart.joined[,set2])
corrplot(tmpcor)
```


## Session Information

```{r session-info}
```
