---
title: "GLMNET Predictions"
author: "Petr Simecek"
output: html_document
---

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

```{r knitr-opts-chunk, include=FALSE}
```

**Last updated:** `r Sys.Date()`

**Code version:** `r workflowr::extract_commit(".", 1)$sha1`


How precisely could we predict age and sex of a mouse from its mRNA / protein expression data? It depends on a tool we use. In this analysis, we utilize [glmnet](https://cran.r-project.org/web/packages/glmnet/index.html) package fitting lasso and elastic-net regularized linear models. 

Each prediction is made without using the animal itself and the quality of glmnet predictions is then judged by comparing those predictions to true values (*jackknife resampling*). 

While sex predictions are very accurate, some animals seem to be predicted younger or older than they really are. Is it a random fluctuation? Or are those signatures of aging shared across different datasets?

The signature of aging seems to be shared within the tissue. For kidney and for heart we observe mRNA and protein based predictions to be correlated.

The between-tissues story is more complicated. On mRNA level, we observe a striking correlation caused by several shared age-outliers (F787, M989, M907, F1208). On protein level we do not see anything like that and after adjusting for age group, the correlation between predictions is not statistically significant. 

The report has been generated from R/Markdown file [predict-glmnet.Rmd](https://github.com/simecek/TheAgingProteome/blob/master/analysis/predict-glmnet.Rmd). Glmnet predictions have been calculated by [predict_glmnet.R](https://github.com/simecek/TheAgingProteome/blob/master/code/predict_glmnet.R) script.

```{r read_data, include=FALSE}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(plotly))
suppressPackageStartupMessages(library(broom))
suppressPackageStartupMessages(library(gridExtra))

pred.kidney <- read.csv("../../kidney2/kidney_predict_glmnet.csv", as.is=TRUE)
pred.heart <- read.csv("../../heart2/heart_predict_glmnet.csv", as.is=TRUE)

# for shared samples, join the predictions into one table
pred.both <- inner_join(pred.kidney, pred.heart, 
                  by = c("Mouse.ID", "Sex", "Age", "Generation"),
                  suffix = c("_kidney", "_heart"))
pred.both$text <- paste0(ifelse(pred.both$Sex==2,"M","F"), pred.both$Sample.Number_kidney)

# function for outputing p-values
pval.output <- function(p) {
  if (p<0.001) 
    return("< 0.001") 
  else 
    paste("=",round(p,3))
}

```


## Quality of Predictions

### Age

```{r qt_age, echo=FALSE}
quality.table <- NULL
cols <- c("Age.prot_kidney", "Age.prot_heart", 
          "Age.mrna_kidney", "Age.mrna_heart")
for (c in cols) {
  new.line <- data.frame(variable = c,
                         rmse = sqrt(mean((pred.both[,c] - pred.both[,"Age"])^2)),
                         cor = cor(pred.both[,c], pred.both[,"Age"]))
  quality.table <- rbind(quality.table, new.line)
}

knitr::kable(quality.table, digits = 2, format = "markdown")
```

```{r quality_kidney_age, echo=FALSE, fig.width=8, fig.height=8}
p1 <- ggplot(pred.both, aes(x=Age, y= Age.mrna_heart, group=Age)) +
  geom_boxplot(outlier.shape=NA) +
  geom_jitter(aes(color=factor(Sex)), width = 0.25) +
  labs(title = "mRNA", subtitle = "Heart") +
  theme(legend.position="none")
p2 <- ggplot(pred.both, aes(x=Age, y= Age.mrna_kidney, group=Age)) +
  geom_boxplot(outlier.shape=NA) +
  geom_jitter(aes(color=factor(Sex)), width = 0.25) +
  labs(title = "mRNA", subtitle = "Kidney") +
  theme(legend.position="none")
p3 <- ggplot(pred.both, aes(x=Age, y= Age.prot_heart, group=Age)) +
  geom_boxplot(outlier.shape=NA) +
  geom_jitter(aes(color=factor(Sex)), width = 0.25) +
  labs(title = "Protein", subtitle = "Heart") +
  theme(legend.position="none")
p4 <- ggplot(pred.both, aes(x=Age, y= Age.prot_kidney, group=Age)) +
  geom_boxplot(outlier.shape=NA) +
  geom_jitter(aes(color=factor(Sex)), width = 0.25) +
  labs(title = "Protein", subtitle = "Kidney") +
  theme(legend.position="none")
grid.arrange(p4, p3, p2, p1, ncol = 2)
```


### Sex 

```{r qt_sex, echo=FALSE}
quality.table <- NULL
cols <- c("Sex.prot_kidney", "Sex.prot_heart", 
          "Sex.mrna_kidney", "Sex.mrna_heart")
for (c in cols) {
  new.line <- data.frame(variable = c,
                         rmse = sqrt(mean((pred.both[,c] - pred.both[,"Sex"])^2)),
                         cor = cor(pred.both[,c], pred.both[,"Sex"]))
  quality.table <- rbind(quality.table, new.line)
}

knitr::kable(quality.table, digits = 2, format = "markdown")
```

```{r quality_kidney_sex, echo=FALSE, fig.width=8, fig.height=8}
p1 <- ggplot(pred.both, aes(x=Sex, y= Sex.mrna_heart, group=Sex)) +
  geom_boxplot(outlier.shape=NA) +
  geom_jitter(aes(color=factor(Age)), width = 0.25) +
  labs(title = "mRNA", subtitle = "Heart") +
  theme(legend.position="none")
p2 <- ggplot(pred.both, aes(x=Sex, y= Sex.mrna_kidney, group=Sex)) +
  geom_boxplot(outlier.shape=NA) +
  geom_jitter(aes(color=factor(Age)), width = 0.25) +
  labs(title = "mRNA", subtitle = "Kidney") +
  theme(legend.position="none")
p3 <- ggplot(pred.both, aes(x=Sex, y= Sex.prot_heart, group=Sex)) +
  geom_boxplot(outlier.shape=NA) +
  geom_jitter(aes(color=factor(Age)), width = 0.25) +
  labs(title = "Protein", subtitle = "Heart") +
  theme(legend.position="none")
p4 <- ggplot(pred.both, aes(x=Sex, y= Sex.prot_kidney, group=Sex)) +
  geom_boxplot(outlier.shape=NA) +
  geom_jitter(aes(color=factor(Age)), width = 0.25) +
  labs(title = "Protein", subtitle = "Kidney") +
  theme(legend.position="none")
grid.arrange(p4, p3, p2, p1, ncol = 2)
```

## Comparison between tissues 

Interactive plots below have titles reporting the comparison group and p-value for test of dependence between predictions (adjusting for age factor).

### Age - mRNA

<center>
```{r sod_age_mrna, echo=FALSE, fig.width=5, fig.height=5}
pval <- subset(tidy(anova(lm(Age.mrna_kidney ~ factor(Age) + Age.mrna_heart, data=pred.both))), term=="Age.mrna_heart")$p.value

pl <- ggplot(pred.both, aes(x=Age.mrna_kidney, y= Age.mrna_heart, color=factor(Age), shape=factor(Sex), text=text)) +
  geom_point() +
  labs(title = paste("Age / mRNA:",  
        "p-value", pval.output(pval))) +
  theme(legend.position="none")
ggplotly(pl, tooltip="text")
```
</center>

### Age - protein

<center>
```{r sod_age_prot, echo=FALSE, fig.width=5, fig.height=5}
pval <- subset(tidy(anova(lm(Age.prot_kidney ~ factor(Age) + Age.prot_heart, data=pred.both))), term=="Age.prot_heart")$p.value

pl <- ggplot(pred.both, aes(x=Age.prot_kidney, y= Age.prot_heart, color=factor(Age), shape=factor(Sex), text=text)) +
  geom_point() +
  labs(title = paste("Age / protein:",
               "p-value", pval.output(pval))) +
  theme(legend.position="none")
ggplotly(pl, tooltip="text")
```
</center>

### Sex - mRNA

<center>
```{r sod_Sex_mrna, echo=FALSE, fig.width=5, fig.height=5}
pval <- subset(tidy(anova(lm(Sex.mrna_kidney ~ factor(Sex) + Sex.mrna_heart, data=pred.both))), term=="Sex.mrna_heart")$p.value

pl <- ggplot(pred.both, aes(x=Sex.mrna_kidney, y= Sex.mrna_heart, color=factor(Sex), shape=factor(Age)), text=text) +
  geom_point() +
  labs(title = paste("Sex / mRNA:", 
              "p-value", pval.output(pval))) +
  theme(legend.position="none")
ggplotly(pl, tooltip="text")
```
</center>

### Sex - protein

<center>
```{r sod_Sex_prot, echo=FALSE, fig.width=5, fig.height=5}
pval <- subset(tidy(anova(lm(Sex.prot_kidney ~ factor(Sex) + Sex.prot_heart, data=pred.both))), term=="Sex.prot_heart")$p.value

pl <- ggplot(pred.both, aes(x=Sex.prot_kidney, y= Sex.prot_heart, color=factor(Sex), shape=factor(Age), text=text)) +
  geom_point() +
  labs(title = paste("Sex / protein:", 
       "p-value", pval.output(pval))) +
  theme(legend.position="none")
ggplotly(pl, tooltip="text")
```
</center>

## Comparison between mRNA and protein levels

### Age - Kidney

```{r sod_age_kidney, echo=FALSE, fig.width=5, fig.height=5}
pval <- subset(tidy(anova(lm(Age.mrna ~ factor(Age) + Age.prot, data=pred.kidney))), term=="Age.prot")$p.value

ggplot(pred.kidney, aes(x=Age.mrna, y= Age.prot, color=factor(Age), shape=factor(Sex))) +
  geom_point(size=I(3)) +
  labs(title = "Age / Kidney", 
       subtitle = paste("p-value", pval.output(pval))) +
  theme(legend.position="none")
```

### Age - Heart

```{r sod_age_heart, echo=FALSE, fig.width=5, fig.height=5}
pval <- subset(tidy(anova(lm(Age.mrna ~ factor(Age) + Age.prot, data=pred.heart))), term=="Age.prot")$p.value

ggplot(pred.heart, aes(x=Age.mrna, y= Age.prot, color=factor(Age), shape=factor(Sex))) +
  geom_point(size=I(3)) +
  labs(title = "Age / Heart", 
       subtitle = paste("p-value", pval.output(pval))) +
  theme(legend.position="none")
```

### Sex - Kidney

```{r sod_sex_kidney, echo=FALSE, fig.width=5, fig.height=5}
pval <- subset(tidy(anova(lm(Sex.mrna ~ factor(Sex) + Sex.prot, data=pred.kidney))), term=="Sex.prot")$p.value

ggplot(pred.kidney, aes(x=Sex.mrna, y= Sex.prot, color=factor(Sex), shape=factor(Age))) +
  geom_point(size=I(3)) +
  labs(title = "Sex / Kidney", 
       subtitle = paste("p-value", pval.output(pval))) +
  theme(legend.position="none")
```

### Sex - Heart

```{r sod_sex_heart, echo=FALSE, fig.width=5, fig.height=5}
pval <- subset(tidy(anova(lm(Sex.mrna ~ factor(Sex) + Sex.prot, data=pred.heart))), term=="Sex.prot")$p.value

ggplot(pred.heart, aes(x=Sex.mrna, y= Sex.prot, color=factor(Sex), shape=factor(Age))) +
  geom_point(size=I(3)) +
  labs(title = "Sex / Heart", 
       subtitle = paste("p-value", pval.output(pval))) +
  theme(legend.position="none")
```

## Session Information

```{r session-info}
```
