---
title: "Histograms of ANOVA p-values"
author: "Petr Simecek"
output: html_document
---

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

```{r knitr-opts-chunk, include=FALSE}
```

**Last updated:** `r Sys.Date()`

**Code version:** `r workflowr::extract_commit(".", 1)$sha1`

For each gene/protein, we have fitted a linear model

$$ y \sim Age + Sex+Generation$$

and then tested significance of Age/Sex effects. The figures below visualise distribution of tests' p-values and compare them between kidney and heart.

The Sex effect is weaker for heart than it is for kidney. For both tissues, the Age effect is much stronger on protein-level that it is on mRNA-level.

The report have been generated from R/Markdown file [pval-historgrams.Rmd](https://github.com/simecek/TheAgingProteome/blob/master/analysis/pval-histograms.Rmd).

## mRNA - Age

```{r mrna_age, echo=FALSE}
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(gridExtra))

# load ANOVA results
aov.kidney <- read.csv("../../kidney2/kidney_anova_table.csv", as.is=TRUE)
aov.heart <- read.csv("../../heart2/heart_anova_table.csv", as.is=TRUE)

p1 <- ggplot(aov.kidney, aes(x=p.mRNA_Age.Sex)) + 
  geom_histogram(binwidth=0.04) +
  theme_bw() +
  labs(title="Kidney", 
       subtitle="mRNA/Age",
       x="p-value")
p2 <- ggplot(aov.heart, aes(x=p.mRNA_Age.Sex)) + 
  geom_histogram(binwidth=0.04) +
  theme_bw() +
  labs(title="Heart", 
       subtitle="mRNA/Age",
       x="p-value")

grid.arrange(p1, p2, ncol = 2)
```

\ 

## mRNA - Sex

```{r mrna_sex, echo=FALSE}
p1 <- ggplot(aov.kidney, aes(x=p.mRNA_Sex.Age)) + 
  geom_histogram(binwidth=0.04) +
  theme_bw() +
  labs(title="Kidney", 
       subtitle="mRNA/Sex",
       x="p-value")
p2 <- ggplot(aov.heart, aes(x=p.mRNA_Sex.Age)) + 
  geom_histogram(binwidth=0.04) +
  theme_bw() +
  labs(title="Heart", 
       subtitle="mRNA/Sex",
       x="p-value")

grid.arrange(p1, p2, ncol = 2)
```

\ 

## Protein - Age

```{r protein_age, echo=FALSE}
p1 <- ggplot(aov.kidney, aes(x=p.Prot_Age.Sex)) + 
  geom_histogram(binwidth=0.04) +
  theme_bw() +
  labs(title="Kidney", 
       subtitle="Protein/Age",
       x="p-value")
p2 <- ggplot(aov.heart, aes(x=p.Prot_Age.Sex)) + 
  geom_histogram(binwidth=0.04) +
  theme_bw() +
  labs(title="Heart", 
       subtitle="Protein/Age",
       x="p-value")

grid.arrange(p1, p2, ncol = 2)
```

\ 

## Protein - Sex

```{r protein_sex, echo=FALSE}
p1 <- ggplot(aov.kidney, aes(x=p.Prot_Sex.Age)) + 
  geom_histogram(binwidth=0.04) +
  theme_bw() +
  labs(title="Kidney", 
       subtitle="Protein/Sex",
       x="p-value")
p2 <- ggplot(aov.heart, aes(x=p.Prot_Sex.Age)) + 
  geom_histogram(binwidth=0.04) +
  theme_bw() +
  labs(title="Heart", 
       subtitle="Protein/Sex",
       x="p-value")

grid.arrange(p1, p2, ncol = 2)
```

\ 

## Session Information

```{r session-info}
```
