---
title: "Heart vs. Kidney Comparison"
author: "Petr Simecek"
output: html_document
---

```{r read-chunk, include=FALSE, cache=FALSE}
knitr::read_chunk("chunks.R")
```

```{r knitr-opts-chunk, include=FALSE}
```

```{r read_data, echo=FALSE}
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(VennDiagram))

# load result tables
dt1 <- read.csv("../../kidney2/kidney_anova_table.csv", as.is=TRUE)
dt2<- read.csv("../../heart2/heart_anova_table.csv", as.is=TRUE) 

# make shared results table
dt <- inner_join(dt1, dt2, by="id", suffix=c("kidney", "heart"))

# adjust all p-values
p.val.cols <- grep("^p[.]", names(dt))
for (i in p.val.cols) {
  dt[,i] <- p.adjust(dt[,i], method="BH")
}

# venn diagram plotting function
my.venn <- function(set.kidney, set.heart, title) {
  venn.diagram( main = title,
                main.cex = sqrt(2),
                x = list(set.kidney , set.heart),
                category.names = c("Kidney" , "Heart"),
                output = TRUE ,
                filename=NULL,
                imagetype="png" ,
                height = 480 , 
                width = 480 , 
                resolution = 300,
                compression = "lzw",
                lwd = 2,
                lty = 'blank',
                fill = c('yellow', 'green'),
                cex = 1,
                fontface = "bold",
                fontfamily = "sans",
                cat.cex = 1.2,
                cat.fontface = "bold",
                main.fontface="bold",
                #  cat.default.pos = "outer",
                cat.pos = c(-27, 27),
                cat.dist = c(0.055, 0.055),
                cat.fontfamily = "sans",
                main.fontfamily = "sans",
                #  rotation = 1
                )
}

pval.output <- function(p) {
  if (p<0.001) 
    return("< 0.001") 
  else 
    paste("=",round(p,3))
}
  
```


**Last updated:** `r Sys.Date()`

**Code version:** `r workflowr::extract_commit(".", 1)$sha1`

There are totally `r nrow(dt)` gene-protein pairs detected both in heart and kidney tissues. 

We used `p.adjust` function (Benjamini & Hochberg method) to adjust ANOVA p-values for Age and Sex effects to multiple testing. Then we used Fisher exact test (as implemented in `fisher.test`) to examine whether age / sex dependent genes / proteins in one tissue are also more likely to be age / sex dependent in the other tissue ($fdr<5\%$).

We conclude that while Age/Sex effects are shared between heart and kidney on mRNA level, those effects seem to be tissue specific for proteins.

The report has been generated from R/Markdown file [pvals-kidney_vs_heart.Rmd](https://github.com/simecek/TheAgingProteome/blob/master/analysis/pvals-kidney_vs_heart.Rmd).

\ 

## mRNA / Age Effect

```{r mrna_age, echo=FALSE}
# mRNA / Age
set1 <- dt$id[dt$p.mRNA_Age.Sexkidney < 0.05]
set2 <- dt$id[dt$p.mRNA_Age.Sexheart < 0.05]
pl <- my.venn(set1, set2, title="")
tab <- table(dt$id %in% set1, dt$id %in% set2)
grid.draw(pl)
#grid.newpage()
```

Genes changing with Age in a heart have significantly higher chance to change also in a kidney and vise verse (Fisher Exact Test, odds ratio = `r round(fisher.test(tab)$estimate,2)`, p-value `r pval.output(fisher.test(tab)$p.value)`).

\ 

## mRNA / Sex Effect

```{r mrna_sex, echo=FALSE}
# mRNA / Age
set1 <- dt$id[dt$p.mRNA_Sex.Agekidney < 0.05]
set2 <- dt$id[dt$p.mRNA_Sex.Ageheart < 0.05]
pl <- my.venn(set1, set2, title="")
tab <- table(dt$id %in% set1, dt$id %in% set2)
grid.draw(pl)
#grid.newpage()
```

Genes changing with Sex in a heart have significantly higher chance to change also in a kidney and vise verse (Fisher Exact Test, odds ratio = `r round(fisher.test(tab)$estimate,2)`, p-value `r pval.output(fisher.test(tab)$p.value)`).

\ 

## Protein / Age Effect

```{r protein_age, echo=FALSE}
# Prot / Age
set1 <- dt$id[dt$p.Prot_Age.Sexkidney < 0.05]
set2 <- dt$id[dt$p.Prot_Age.Sexheart < 0.05]
pl <- my.venn(set1, set2, title="")
tab <- table(dt$id %in% set1, dt$id %in% set2)
grid.draw(pl)
#grid.newpage()
```

There seems to be no statistically significant relation between whether a protein is correlated to Age in heart and kidney (Fisher Exact Test, odds ratio = `r round(fisher.test(tab)$estimate,2)`, p-value `r pval.output(fisher.test(tab)$p.value)`).

\ 

## Prot / Sex Effect

```{r prot_sex, echo=FALSE}
# mRNA / Age
set1 <- dt$id[dt$p.Prot_Sex.Agekidney < 0.05]
set2 <- dt$id[dt$p.Prot_Sex.Ageheart < 0.05]
pl <- my.venn(set1, set2, title="")
tab <- table(dt$id %in% set1, dt$id %in% set2)
grid.draw(pl)
#grid.newpage()
```

There seems to be no statistically significant relation between whether a protein is correlated to Sex in heart and kidney (Fisher Exact Test, odds ratio = `r round(fisher.test(tab)$estimate,2)`, p-value `r pval.output(fisher.test(tab)$p.value)`).

## Session Information

```{r session-info}
```
